### 2025-04-27 paper/paper.tex
Update paper.tex

A third research gap appears to be the influence of various real-world distractions that can occur, ranging from road accidents, pedestrians, and emergency vehicles. Although Du et al. have begun to account for pedestrians \cite{13}, there are still several factors that go unaccounted for in the process of traffic signal prediction, which can back up traffic worse than threshold options.

@@

After beginning implementation we decided that CityFlow was not the appropriate tool for this problem, despite our research. While a powerful tool, it struggled to handle the data format provided from OSM data, resulting in slow performance and broken intersections. We then decided moving to the SUMO simulator was a more worthwhile decision, as its efficiency has improved over time and documentation is more appropriate for our specific research area.

Once SUMO was installed, we downloaded public data for the city of Starkville Mississippi's road structure using the tool open street map, and used SUMO's netedit to process the data into a usable format. Some changes were made, such as fixing traffic signal placement using junctions, and fixing non-signal based intersection interactions.

@@

Using the setup configuration file, we can use SUMO GUI, which comes bundled with the installation package of SUMO. This tools allows us to run a visualization of the simulation with regards to the currently implemented method of traffic signal control, and make observations from the point of view of drivers without quantitative analysis, making observations about our model to make improvements. a

@@

The next is the average trip time. Because we will benchmark similar environments, we can get the average trip times of all combined vehicles in a simulation, and use it as a quantitative performance metric. Quicker trips indicates a successful traffic signal control implementation, as people can get to their end destination quicker, the entire point of driving. It also serves as a benchmark for other metrics, likely indicating shorter queue lengths, less stops, and less time waiting at stops on average.

@@

our neural network is implemented through the python libraries traffic control interface (traci) and pytorch. Traci allows us to modify our simulation and get state information within a current timestep, while pytorch allows us to easily build and implement neural networks for our project.

The first step in designing the proposed neural network is to establish what an agent is, and how they will communicate with the model. Based on existing codebases and research, we decided implementing a decentralized multi agent reinforcement learning model. This specific implementation lacks a central coordinator, and each agent acts on its own in coordination with other agents. The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time; it also accepted a list of possible actions the traffic signal controller could perform, which when provided to a deep-q learning model can guess the optimal move.

@@

Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and timeframe would not allow us to do this. Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly backed up. Most of these points were decided by considering the importance of the road, the time spent at intersections, and personal experience. This way we don't waste computational resources at intersections that see little traffic, or do fine directing traffic on their own. Doing so significantly decreases the computational requirements to run the model, and also allows us to focus only on important points.

>>>>

A third research gap appears to be the influence of various real-world distractions that can occur, ranging from road accidents, pedestrians, and emergency vehicles. Although Du et al. have begun to take into account pedestrians \cite{13}, there are still several factors that remain unaccounted for in the traffic signal prediction process, which can back up traffic worse than threshold options.

@@

After starting implementation, we decided that CityFlow was not the appropriate tool for this problem, despite our research. While a powerful tool, it struggled to handle the data format provided from OSM data, resulting in slow performance and broken intersections. We then decided moving to the SUMO simulator was a more worthwhile decision, as its efficiency has improved over time and documentation is more appropriate for our specific research area.

Once SUMO was installed, we downloaded public data for the city of Starkville Mississippi's road structure using the tool open street map, and used SUMO's netedit to process the data into a usable format. Some changes were made, such as fixing traffic signal placement using junctions and fixing non-signal-based intersection interactions.

@@

Using the configuration file, we can use the SUMO GUI, which comes included with the SUMO installation package. This tools allows us to run a visualization of the simulation with regards to the currently implemented method of traffic signal control, and make observations from the point of view of drivers without quantitative analysis, making observations about our model to make improvements. a

@@

The next is the average trip time. Because we will benchmark similar environments, we can get the average trip times of all combined vehicles in a simulation, and use it as a quantitative performance metric. Faster trips indicate a successful implementation of traffic signal control, as people can get to their end destination faster. It also serves as a benchmark for other metrics, likely indicating shorter queue lengths, less stops, and less time waiting at stops on average.

@@

Our neural network is implemented through the python libraries traffic control interface (traci) and pytorch. Traci allows us to modify our simulation and get state information within a current timestep, while pytorch allows us to easily build and implement neural networks for our project.

The first step in designing the proposed neural network is to establish what an agent is, and how they will communicate with the model. Based on existing codebases and research, we decided to implement a decentralized multi-agent reinforcement learning model. This specific implementation lacks a central coordinator, and each agent acts on its own in coordination with other agents. The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time; it also accepted a list of possible actions the traffic signal controller could perform, which when provided to a deep-q learning model can guess the optimal move.

@@

Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and time frame would not allow us to do this. Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly backed up. Most of these points were decided by considering the importance of the road, the time spent at intersections, and personal experience. This way we don't waste computational resources at intersections that see little traffic, or do fine directing traffic on their own. Doing so significantly decreases the computational requirements to run the model, and also allows us to focus only on important points.
